---
title: "SF Demo"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

To prevent any migraines, it's best to use R within RStudio and use separately from Python. Although you can use both within the same notebook/RMarkdown, its kinda annoying.

- Use install.packages('sf') to install

It's also useful to have tidyverse installed in case you want to do some data wrangling before mapping your spatial data.

```{r}
library(tidyverse)
library(sf)
library(leaflet)
library(mapview)
```

## Reading in data
```{r}
buildings_sdf = st_read("data/buildings_pts.shp")
tibble(buildings_sdf)
```

## Static Mapping
```{r}
ggplot(buildings_sdf)+
  geom_sf(aes(color = btype), size = 2) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

# Buffer Example
```{r}
bus_lanes_sdf = st_read('data/wb_LocalBus.shp')
tibble(bus_lanes_sdf)
```

```{r}
# check the coordinate reference system units (meters)
st_crs(bus_lanes_sdf)
```

```{r}
ggplot(bus_lanes_sdf)+
  geom_sf(aes(color =v832_SE_37), size = 2) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
# Filter buslanes to Weber COunt and plot
bus_lanes_weber <- bus_lanes_sdf %>%
  filter(v832_SE_37 == 'WEBER')

ggplot(bus_lanes_weber) +
  geom_sf(aes(color = 'red')) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  theme(legend.position="none")
```

```{r}
meters_to_buffer <-  .4 * 1609.344
buffered_lanes = st_buffer(bus_lanes_weber, meters_to_buffer)
```

```{r}
ggplot() +
  geom_sf(data = buffered_lanes, aes(color = 'red')) +
  geom_sf(data = bus_lanes_weber, aes(color = 'blue')) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  theme(legend.position="none")
```

# Spatial Join Examples
```{r}
weber_cities_sdf = st_read('data/weber_cities.shp')
ggplot(weber_cities_sdf) +
  geom_sf(aes(fill = NAME)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
# Join Cities attributes to the buildings points
cities_buildings_join = st_join(weber_cities_sdf, buildings_sdf, join = st_intersects)
```

```{r}
cb_agg = cities_buildings_join %>%
  group_by(NAME) %>%
  summarize(res_units = sum(res_units), job_spaces = sum(job_spaces))
tibble(cb_agg)
```
```{r}
cb_centroids = st_centroid(cb_agg)
cb_centroids_coords = st_coordinates(cb_centroids)
cb_centroids2 = cb_centroids %>%
  mutate(x = cb_centroids_coords[,1], y = cb_centroids_coords[,2])
ggplot(cb_agg) +
  geom_sf(aes(fill = NAME)) +
  geom_text(data = cb_centroids2, aes(x, y, label = NAME), colour = "white", size = 2) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
## TDM Example: Join Link Midpoints with Toll zones with Spatial join

```{r}
# read in data
out_link_mp <- st_read('data/C1_Link_Midponts.shp')
tollz_shp <- st_read('data/Tollz_shp.shp')
```

```{r}
mp_tollz = st_join(out_link_mp, tollz_shp, join = st_within, left = FALSE)
tibble(mp_tollz)
```
```{r}
# write the result to shapefile
#st_write(mp_tollz, 'Outputs/r_mp_tollz.shp')
```

# Interactive Maps
```{r}
weber_cities_leaflet <- weber_cities_sdf %>%
  st_transform(4326) %>%
  mutate(col=sample(c('red','yellow','green','blue','orange','brown','purple'),nrow(weber_cities_sdf),1))
```

```{r}
leaflet(weber_cities_leaflet) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(
    data = weber_cities_leaflet,
    co = "black", 
    fillColor = weber_cities_leaflet$col,
    highlightOptions = highlightOptions( color='red', weight=4, bringToFront = TRUE),
    popup = paste0(
      "<b>NAME: </b>", weber_cities_leaflet$NAME, "<br>",
      "<b>AREA: <b/>", weber_cities_leaflet$AREA, "<br>"
  ))
```


```{r}
mapview(weber_cities_leaflet, zcol = "NAME")
```








